---
title: "Basic programming"
author: "Stewart Li"
date: "8/29/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```


# Document structure


1. R || Python || Javascript
2. Base R || Pandas
3. tidyverse and data.table


# Variables


Variable types: character, numeric, integer, double, logical, date, factor, complex, raw, NULL.


```{r}
class(c("strings"))
typeof(c(1L))
typeof(c(1))
typeof(c(TRUE))
class(as.Date(Sys.time()))
class(factor(c("hello", "hi", "bye", "bye")))
typeof(c(1i))
class(c(charToRaw("strings")))
```


`is` and `as`. 


```{r}
is.double(c(2))
is.nan(c(-2, 0, 2) / 0)
is.vector(numeric(10))

class(toString(c(2)))
class(as.numeric(c("2")))
```


Objects assignment: `=`, `<-`, `->`, `<<-`, 


```{r}
my.name <- readline(prompt="Enter name: ") # input in python
print(paste("Hi,", my.name))
```


# Data structure


Vector: same data type, recycled, vectorised.
Index and names: `$, [], [[]], .[]`. 


```{r}
c(11, "you", TRUE ) # without losing information
c(11, TRUE )

c(c(1:3), c(1:10))
c(2, 4, 6)^2

intersect(c(3:4), c(1:10))
union(c(11, 12), c(1:10))

numeric(10) # create a vector
character(10)
```


```{r}
?`[` # i, j, k

a <- c(you = 3, me = 4)
names(a)
tracemem(a) # id

a['you'] # like counter in python
a['them'] # NA
a[1:2] # slice 
```


Matrices: same data type.
Array: multiple dimensions.


```{r}
matrix(letters, 13, 2, byrow = TRUE) 

solve(matrix(1:26, 2, 2)) # inverse
matrix(1:26, 2, 2) %*% solve(matrix(1:26, 2, 2))
diag(2)

t(head(mtcars))
```


List: anything.
Dataframe: same length, rows and columns.


```{r}
help("data.frame")

data.frame(id = 1:6,
           greet = sample(LETTERS, 6),
           nums = runif(6, 0, 100)) 
```


```{r}
lista <- list(a = 1:10, 
              b = c("you", "and", "me"), 
              c = c(TRUE, FALSE), 
              d = head(mtcars)) # like dict in python

lista[1] # extract
lista["a"]
lista[[1]]
lista$a

lista$e <- c(20:26) # create 
append(lista, c(20:26)) # add 

lista[5] <- NULL # remove 
lista[["a"]] <- c(13:19) # modify
```


# Operators


Arithmetic calculator: `+, -, *, /, ^, %/%, %%`.
Others: `;`, `#`, 


```{r}
3 + 6
`+` (3 + 6)

49 %/% 2
```


# Conditions


Logical operators: `==, !=, >, >=, |, &, !, %in%, -, xor`, 
Control flow: `if`, `switch`, 


```{r}
"you" == "you" 
"you" != "me"
```


```{r}
if(49 > 49)
  {print('More')
}else if(49 > 50 | 0 >= -1){
    print("Less")
}else{
    print("None")
  }  
```


```{r}
for (i in 1:10) {
  if(i %% 2 == 0){
    print(paste0("The number ", i, " is divisible by 2."))
  }else if(i %% 3 == 0){
    print(paste0("The number ", i, " is divisible by 3."))
  }else{
    print(paste0("The number ", i, " is not divisible by 2 or 3."))
  }
}
```


# Loop


```{r}
for (i in 1:10) {
  print(i^2) # reset i each time
}

for (i in letters) {
  print(i)
}

for (i in seq_along(1:10)){
  print(paste0("this is ", i)) # index vs items), length(), seq_along()
} 

ls() # last object
```


Iterate over index or item.  


```{r}
a <- c(2, 5, "you")

for (x in a){ 
  if(x == "5"){
    print("you are so cool")
  }
} 

for (x in 1:length(a)){  
  if(a[x] == "5"){
    print("you are so cool")
  }
}
```


```{r}
i <- 1:6
for (i in i) {
  if (i == 3){
    next
  }
  print(i)
}

i <- 1
while (i < 6) {
  if (i == 3){
    break
  }
  print(i)
  i = i+1
}
```


Pre-allocation of object. 


```{r}
a <- vector(mode = 'list', length = 10)

for (i in 1:10) {
  a[i] <- i^2
}

unlist(a)
```


```{r}
a <- (1:10)
for (x in a){
  print(x^x)
}

c(1:10) %>% map_dbl(~.x^.x)
```


# Functions


```{r}
myfun <- function(x, y = 10){
  a <- x + y
  a <- paste0('result is a number of ', a)
  a
} 

myfun(20)
```


```{r}
myfun1 <- function(func, ...){
  func(mtcars$mpg)
} # input function

myfun1(mean)
myfun1(sum)
```


```{r}
myfun2 <- function(x){
  addmy <- function(){
    paste0("start ", mean(c(x, 10)), " end")
  }
  return(addmy)
} # output function

myfun2(90)()
```


# Erorr handling


```{r}
myupdate <- function(x){
    tryCatch(
      {
        y = x * 2
        return(y)
      },
      error = function(error_message) {
        message("This is my custom message.")
        cli::cli_alert_danger("what are you doing?") 
        message("And below is the error message from R:")
        return(message(error_message)) # catch errors***
      }
    )
}

myupdate("a")
```


# OOP


```{r}
Audit <- R6::R6Class(
  classname = "Audit",
  public = list(
    initialize = function(client, aic) {
      private$client <- client
      private$aic <- aic
      # if (!missing(client)) self$client <- client
      # if (!missing(aic)) self$aic <- aic
      self$greet()
    },
    set_aic = function(val) {
      private$aic <- val
    },
    greet = function() {
      cat(paste0(private$client, " is charge of the audit of ", private$aic, ".\n"))
    }
  ), 
  private = list(
    client = NA,
    aic = NA
  )
)

nh2020 <- Audit$new("RAudit Solution LLP", "Stewart Li")

nh2020$set_aic("Song Peng")
nh2020$greet()
```


# Data sciense


1. File type: `.R`, `.Rmd`, `Rdata`, 
2. Import library: `dplyr::select`, `:::`, 


```{r}
getwd()
dir(".") 
list.files() 
```


```{r}
?mtcars

attributes(mtcars)
dim(mtcars)
str(mtcars)
head(mtcars);tail(mtcars)

Map(min, mtcars)
apply(mtcars, 2, min)
Reduce(`+`, mtcars)

with(mtcars, by(am, gear, summary))
with(mtcars, aggregate(mpg, by = list(am, gear), FUN = function(x) c(mean(x), sd(x))))

mean(mtcars$mpg)
mean(mtcars[mtcars$hp > mean(mtcars$hp), "mpg"])

transform(mtcars, e = hp*.01)
merge(mtcars, data.frame(am = c("yes", "no")), by = "am", all = TRUE) # row bind
```


```{r}
plot(mtcars) # pairs()
hist(mtcars$mpg)

plot(mpg ~ hp, 
     data = mtcars, 
     main = 'Scattor plot', 
     xlab = 'MPG', 
     ylab = 'HP', 
     cex.main = 1.5, 
     cex.lab = 1.5, 
     col = 'red', 
     pch = 16)
abline(h = mean(mtcars$mpg), col = 'red')
abline(v = mean(mtcars$hp), col = 'blue')
```


```{r}
pagedown::chrome_print(here::here("tutorials/1_programming_r.html"), here::here("tutorials/1_programming_r.pdf"))
```





