---
title: "Data Squad Technical Test"
author: "Stewart Li"
date: "4/7/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, error = FALSE, comment = '>')

library(tidyverse)
library(lubridate)
library(openxlsx)
```


*Source:* [*Data Squad Technical Test*](https://github.com/cdg-data-coe/interviews2)


## Context

Under each transport commercial contract, there may be multiple, dynamic bookings per month. There is opportunity to automate and optimise transportation planning i.e. match vehicles to schedule.     

The goal is to allocate the minimum number of vehicles, (and cost - larger cacpacity bus cost more) to each commercial contract for that 1 month.     

Each trip journey, and each bus total journey **cannot exceed 45mins**. There are 10, 19 and 40-seater buses that can be utilised.    


### Task

1. Perform a EDA on the excel sheet named- 'input' using MS Power BI.    
2. Then, write a Python program to generate the 'output_plan' excel sheet.    
3. Ensure the code / model can run on other scenario inputs with same data attributes.    
4. State your assumptions clearly.    


### Important Instructions

1. Implement the task, comment your codes and commit into github.    
2. Write clear steps for evaluators to setup and run your codes.    
3. Include comments on how your solution solves the problem and possible enhancement if given more time and data (what kind of data).    


## My solution


Assumptions;    

1. I am unable to find information about the column of `10, 19 and 40-seater buses`. Therefore, I can not do calculation (or model) on it.     
2. Columns of `pickup_time` and `dropoff_time` contain missing values. I remove them in order to calculate travel time.     
3. The columns of `bookiing_id` is unique (no duplication), which seems different from the question statement (The goal is to allocate the minimum number of vehicles to each commercial contract for that 1 month). It would be helpful if there are more details in this regards.  


Notes;    

1. I use `R` to perform EDA instead of `MS Power BI`.   
2. It is very easy to write out excel sheet in `R`. Thus, it is not necessary for me to jump into `Python` to do it.  
3. My result is the `stewartli` (sheetname) of excel workbook. HTML version of this document is attached for your review.   
4. I have refactored my code to ensure it runs for other similar data.   


### Import data


```{r}
# step1: read in data
df <- readxl::read_excel('./Bus-trips-data-sample.xlsx')

# view data structure (like Python Pandas describe function)
glimpse(df)
```


### Clean data


```{r}
# step2: data cleaning
df_clean <- df %>% 
  # clean column names
  janitor::clean_names() %>% 
  # convert to date
  mutate(pickup = paste(journey_date, pickup_time, sep = " "), 
         dropoff = paste(journey_date, dropoff_time, sep = " "), 
         journey = dmy(journey_date)) %>%
  # convert missing values to NA
  mutate(pickup_time = ifelse(pickup_time == "--:--", NA, pickup_time), 
         dropoff_time = ifelse(dropoff_time == "--:--", NA, dropoff_time)) %>% 
  # select relevant columns
  select(booking_id, journey, pickup, dropoff, everything()) 

# view first 6 rows of dataset
head(df_clean)
```


### EDA


```{r}
# step3: overview of the dataset
skimr::skim(df_clean)

# visual overview of the dataset 
visdat::vis_dat(df_clean)
```


```{r}
# step4: calculate cancel rate of weekdays
df_clean %>% 
  # create a weekday column
  mutate(weekday = wday(journey, label = TRUE)) %>%
  # calculate number of booking_id based on weekday and booking_status
  group_by(weekday, booking_status) %>% 
  summarise(n = n(), .groups = "drop") %>% 
  # reshape the dataset
  pivot_wider(names_from = booking_status, values_from = n, values_fill = 0) %>% 
  # calculate cancel rate of weekdays
  mutate(pct_cancel = Cancelled / (Cancelled + Confirmed + New)) 
```


### Write excel


```{r}
# step5: prepare the required data structure
travel <- df_clean %>% 
  # filter out NA
  dplyr::filter(!is.na(pickup_time), !is.na(dropoff_time)) %>% 
  # calculate the travel time
  mutate(pickup1 = str_remove(pickup, " AM"), 
         pickup1 = parse_datetime(pickup1, "%d %b %Y %H:%M"), 
         dropoff1 = str_remove(dropoff, " AM"), 
         dropoff1 = parse_datetime(dropoff1, "%d %b %Y %H:%M"), 
         travel = dropoff1 - pickup1) %>% 
  # filter out travel time exceeding 45 mins
  dplyr::filter(travel < 45) %>% 
  # select relevant columns
  select(journey, travel, booking_id) %>% 
  # paste booking_id based on journey_date and travel time
  group_by(journey, travel) %>% 
  summarise(booking_ids = paste(booking_id, collapse = ', '), .groups = "drop") %>% 
  # create this column
  mutate(type_of_bus_10_19_40_seaters = 10) %>% 
  # get the required data shape
  select(journey, type_of_bus_10_19_40_seaters, travel, booking_ids)

# view first 6 rows of dataset
head(travel)
```


```{r eval=FALSE}
# step6: write out excel
wb <- loadWorkbook("Bus-trips-data-sample.xlsx")
addWorksheet(wb, "stewartli")
writeData(wb, "stewartli", travel, startRow = 1, startCol = 1)
saveWorkbook(wb, file = "Bus-trips-data-sample1.xlsx")
```


```{r}
knitr::knit_exit()
```






